---
title: "Implementing Monopoly in R"
author: "Noah Jones"
date: "10/12/2021"
output: pdf_document
---

# Using R6 library to implement object-oriented classes and class specific functions for a traditional Monopoly game, followed by a variety of test cases and a simulation of 1000 games.

```{r}
library(R6)

# gameboard and decks 

gameboard <- data.frame(
  space = 1:40, 
  title = c(
    "Go", "Mediterranean Avenue", "Community Chest", "Baltic Avenue",
    "Income Tax", "Reading Railroad", "Oriental Avenue", "Chance",
    "Vermont Avenue", "Connecticut Avenue", "Jail", "St. Charles Place",
    "Electric Company", "States Avenue", "Virginia Avenue",
    "Pennsylvania Railroad", "St. James Place", "Community Chest",
    "Tennessee Avenue", "New York Avenue", "Free Parking",
    "Kentucky Avenue", "Chance", "Indiana Avenue", "Illinois Avenue",
    "B & O Railroad", "Atlantic Avenue", "Ventnor Avenue", "Water Works",
    "Marvin Gardens", "Go to jail", "Pacific Avenue",
    "North Carolina Avenue", "Community Chest", "Pennsylvania Avenue",
    "Short Line Railroad", "Chance", "Park Place", "Luxury Tax",
    "Boardwalk"), stringsAsFactors = FALSE)
chancedeck <- data.frame(
  index = 1:15, 
  card = c(
    "Advance to Go", "Advance to Illinois Ave.",
    "Advance to St. Charles Place", "Advance token to nearest Utility",
    "Advance token to the nearest Railroad",
    "Take a ride on the Reading Railroad",
    "Take a walk on the Boardwalk", "Go to Jail", "Go Back 3 Spaces",
    "Bank pays you dividend of $50", "Get out of Jail Free",
    "Make general repairs on all your property", "Pay poor tax of $15",
    "You have been elected Chairman of the Board", 
    "Your building loan matures"), stringsAsFactors = FALSE)
communitydeck <- data.frame(
  index = 1:16, 
  card = c(
    "Advance to Go", "Go to Jail",
    "Bank error in your favor. Collect $200", "Doctor's fees Pay $50",
    "From sale of stock you get $45", "Get Out of Jail Free",
    "Grand Opera Night Opening", "Xmas Fund matures", "Income tax refund",
    "Life insurance matures. Collect $100", "Pay hospital fees of $100",
    "Pay school tax of $150", "Receive for services $25",
    "You are assessed for street repairs",
    "You have won second prize in a beauty contest",
    "You inherit $100"), stringsAsFactors = FALSE)

# RandomDice class 

RandomDice <- R6Class(
  classname = "RandomDice",
  public = list(
    verbose = NA,
    initialize = function(verbose = FALSE){
      stopifnot(is.logical(verbose))
      self$verbose = verbose
    },
    roll = function() {
      outcome <- sample(1:6, size = 2, replace = TRUE)
      if(self$verbose){
        cat("Dice Rolled:", outcome[1], outcome[2], "\n")
      }
      outcome
    }
  )
)

# Preset Dice 

PresetDice <- R6Class(
  classname = "PresetDice",
  public = list(
    verbose = NA,
    preset_rolls = double(0),
    position = 1,
    initialize = function(rolls, verbose = FALSE){
      stopifnot(is.logical(verbose))
      stopifnot(is.numeric(rolls))
      self$preset_rolls = rolls
      self$verbose = verbose
    },
    roll = function(){
      if(self$position > length(self$preset_rolls)){
        stop("You have run out of predetermined dice outcomes.")
      }
      outcome <- c(self$preset_rolls[self$position], 
                   self$preset_rolls[self$position + 1])
      self$position <- self$position + 2
      if(self$verbose){
        cat("Dice Rolled:", outcome[1], outcome[2], "\n")
      }
      outcome
    }
  )
)


# Chance and Community Decks 

# This R6 class object shuffles the card deck when initialized.
# It has one method $draw(), which will draw a card from the deck.
# If all the cards have been drawn (position = deck length), then it will
# shuffle the cards again.
# The verbose option cats the card that is drawn on to the screen.
CardDeck <- R6Class(
  classname = "CardDeck",
  public = list(
    verbose = NA,
    deck_order = double(0), 
    deck = data.frame(),
    position = 1,
    initialize = function(deck, verbose = FALSE){
      stopifnot(is.data.frame(deck),
                is.numeric(deck[[1]]),
                is.character(deck[[2]]))
      self$deck_order <- sample(length(deck[[1]]))
      self$verbose <- verbose
      self$deck <- deck
    },
    draw = function(){
      if(self$position > length(self$deck_order)){
        # if we run out of cards, shuffle deck
        # and reset the position to 1
        if(self$verbose){
          cat("Shuffling deck.\n")
        }
        self$deck_order <- sample(length(self$deck[[1]]))
        self$position <- 1
      }
      outcome <- c(self$deck_order[self$position]) # outcome is the value at position
      self$position <- self$position + 1 # advance the position by 1
      if(self$verbose){
        cat("Card:", self$deck[outcome, 2], "\n")
      }
      outcome # return the outcome
    }
  )
)


# R6 Class SpaceTracker 

SpaceTracker <- R6Class(
  classname = "SpaceTracker",
  public = list(
    counts = rep(0, 40),
    verbose = TRUE,
    tally = function(x){
      self$counts[x] <- self$counts[x] + 1
      if(self$verbose){
        cat("Added tally to ", x, ": ", gameboard$title[x], ".\n", sep = "")
      }
    },
    initialize = function(verbose){
      self$verbose <- verbose
    }
  )
)

# R6 Class Player 

Player <- R6Class(
  classname = "Player",
  public = list(
    pos = 1,
    verbose = TRUE,
    jail = FALSE,
    jail_num = 0,
    move_fwd = function(n){
      self$pos <- self$pos + n
      if(self$pos > 40){
        self$pos <- self$pos - 40
      }
      if(self$verbose){
        cat("Player is now at ", self$pos, ": " , gameboard$title[self$pos], "\n", sep = "")
      }
    },
    initialize = function(verbose = FALSE, pos = 1, jail = FALSE, jail_num = 0) {
      self$verbose <- verbose
      self$pos <- pos
      self$jail <- jail
      self$jail_num <- jail_num
    }
  )
)


# turn taking example 

community_chest_result <- function(n) {
  if(n == 1) {
    return(1)
  } else if(n == 2) {
    return(11)
  } else {
    return(0)
  }
}
chance_result <- function(n, player_rolling) {
  if(n == 1) {
    return(1)
  } else if(n == 2) {
    return(25)
  } else if(n == 3) {
    return(12)
  } else if(n == 4) {
    current <- player_rolling$pos
    if(current == 23) {
      return(29)
    } else {
      return(13)
    }
  } else if(n == 5) {
    current <- player_rolling$pos
    if(current == 8) {
      return(16)
    } else if(current == 23) {
      return(26)
    } else {
      return(6)
    }
  } else if(n == 6) {
    return(6)
  } else if(n == 7) {
    return(40)
  } else if(n == 8) {
    return(11)
  } else if(n == 9) {
    return(player_rolling$pos - 3)
  } else {
    return(0)
  }
}



take_turn <- function(player, spacetracker){
    doubles <- TRUE
    num_doubles <- 0
    while(doubles) {
      dice_rolls <- dice$roll()
      if(player$jail) {
        if(dice_rolls[1] == dice_rolls[2]) {
          if(player$verbose) {
            cat("In jail but rolled doubles.", "\n")
            cat("Player exits jail.", "\n")
          }
          player$jail_num <- 0
          doubles <- FALSE
          player$jail <- FALSE
          player$jail_num <- 0
        } else if(player$jail_num == 2){
          if(player$verbose) {
            cat("Player's third turn in jail. Player must exit jail.", "\n")
          }
          doubles <- FALSE
          player$jail <- FALSE
          player$jail_num <- 0
        } else {
          if(player$verbose) {
            cat("Player stays in jail.", "\n")
          }
          player$jail_num <- player$jail_num + 1
          doubles <- FALSE
          spacetracker$tally(player$pos)
          break
        } 
      } else if(dice_rolls[1] != dice_rolls[2]) {
        doubles <- FALSE
      } else {
        num_doubles <- num_doubles + 1
        if(player$verbose) {
          cat("Doubles count is now", num_doubles, "\n")
        }
      }
      if(num_doubles == 3) {
        player$pos <- 11
        player$jail <- TRUE
        if(player$verbose) {
          cat("Player goes to Jail.", "\n")
        }
        spacetracker$tally(player$pos)
        break
      }
      if(player$verbose) {
        cat("Player starts at ", player$pos, ": ", gameboard$title[player$pos], "\n", sep = "")
      }
      player$move_fwd(sum(dice_rolls))
      if(player$pos == 31) {
        if(player$verbose) {
          cat("Player goes to jail.", "\n")
        }
        player$pos <- 11
        player$jail <- TRUE
        spacetracker$tally(player$pos)
        break
      }
      if(player$pos %in% c(3, 18, 34) ) {
        num <- community$draw()
        result <- community_chest_result(num)
        if(result != 0) {
          if(result == 11) {
            spacetracker$tally(player$pos)
            if(player$verbose) {
              cat("Player goes to jail.", "\n")
            }
            player$pos <- 11
            player$jail <- TRUE
            spacetracker$tally(player$pos)
            break
          } else {
            spacetracker$tally(player$pos)
            player$move_fwd(40 + result - player$pos)
          }
        }
      }
      if(player$pos %in% c(8, 23, 37) ) {
        num <- chance$draw()
        result <- chance_result(num, player)
        if(result != 0) {
          if(result == 11) {
            spacetracker$tally(player$pos)
            if(player$verbose) {
              cat("Player goes to jail.", "\n")
            }
            player$pos <- 11
            player$jail <- TRUE
            spacetracker$tally(player$pos)
            break
          } else {
            spacetracker$tally(player$pos)
            player$move_fwd(40 + result - player$pos)
          }
        }
      }
      spacetracker$tally(player$pos)
      if(doubles && player$verbose) {
        cat("\n", "Player rolled doubles, so they take another turn.", "\n", sep = "")
      }
    }
  }



```


\pagebreak

# Part 1: Test Cases

## do not alter the code for the test cases

### Test Case 1: Space: Go to Jail

```{r, error = TRUE}
dice <- PresetDice$new(
  rolls = c(3,4),
  verbose = TRUE
)
set.seed(16)
player1 <- Player$new(verbose = TRUE, pos = 24)
monopoly <- SpaceTracker$new(verbose = TRUE)

for(i in 1:1){
  cat("--- Turn", i,"---\n")
  take_turn(player1, monopoly)
  cat("\n")
}

print(setNames(monopoly$counts, 1:40))
```

\pagebreak

### Test Case 2: Chance Card and Doubles Tests: Advance to Go, Reading Railroad, Nearest Railroad, Nearest Utility, No Movement

```{r, error = TRUE}
dice <- PresetDice$new(
  rolls = c(3,4, 4,3, 1,1, 3,4, 5,3),
  verbose = TRUE
)
RNGkind(sample.kind = "Rejection")
set.seed(135)
chance <- CardDeck$new(chancedeck, verbose = TRUE)
community <- CardDeck$new(communitydeck, verbose = TRUE)
player1 <- Player$new(verbose = TRUE)
monopoly <- SpaceTracker$new(verbose = TRUE)

for(i in 1:4){
  cat("--- Turn", i,"---\n")
  take_turn(player1, monopoly)
  cat("\n")
}

print(setNames(monopoly$counts, 1:40))
```
\pagebreak

### Test Case 3: Multiple doubles. Community Chest.
```{r, error = TRUE}
dice <- PresetDice$new(
  rolls = c(3,3, 2,2, 2,1, 3,1), verbose = TRUE)

player1 <- Player$new(verbose = TRUE)
monopoly <- SpaceTracker$new(verbose = TRUE)
for(i in 1:2){
  cat("--- Turn", i,"---\n")
  take_turn(player1, monopoly)
  cat("\n")
}

print(setNames(monopoly$counts, 1:40))
```

\pagebreak

### Test Case 4: Doubles three times. Three turns in jail.
```{r, error = TRUE}
dice <- PresetDice$new(
  rolls = c(3,3, 3,3, 3,3, 5,6, 5,6, 5,6),
  verbose = TRUE
)

player1 <- Player$new(verbose = TRUE)
monopoly <- SpaceTracker$new(verbose = TRUE)

for(i in 1:4){
  cat("--- Turn", i,"---\n")
  take_turn(player1, monopoly)
  cat("\n")
}

print(setNames(monopoly$counts, 1:40))
```

\pagebreak

### Test Case 5: After going to Jail, player's turn ends immediately. Rolling doubles while in Jail gets player out of jail.
```{r, error = TRUE}
dice <- PresetDice$new(
  rolls = c(3,3, 1,2, 3,3, 3,4),
  verbose = TRUE
)

player1 <- Player$new(verbose = TRUE, pos = 25)
monopoly <- SpaceTracker$new(verbose = TRUE)

for(i in 1:3){
  cat("--- Turn", i,"---\n")
  take_turn(player1, monopoly)
  cat("\n")
}

print(setNames(monopoly$counts, 1:40))
```

\pagebreak

### Test Case 6: 20 Predetermined Turns

```{r, error = TRUE}
## You must use these dice for Part 1
dice <- PresetDice$new(
  rolls = c(6,4, 5,3, 3,5, 4,4, 4,4, 2,2, 4,3, 4,4, 1,4,
            3,4, 1,2, 3,6, 5,4, 5,5, 1,2, 5,4, 3,3, 6,1,
            1,1, 2,3, 5,5, 5,4, 4,1, 2,2, 2,4),
  verbose = TRUE
)
set.seed(2)
chance <- CardDeck$new(chancedeck, verbose = TRUE)
community <- CardDeck$new(communitydeck, verbose = TRUE)
```


```{r, error = TRUE}
player1 <- Player$new(verbose = TRUE)
monopoly <- SpaceTracker$new(verbose = TRUE)

for(i in 1:20){
  cat("--- Turn", i,"---\n")
  take_turn(player1, monopoly)
  cat("\n")
}

monopoly$counts
cbind(gameboard, counts = monopoly$counts)
```


# Part 2: 1000 simulated games
```{r, message = FALSE}
library(dplyr)
```


```{r, error = TRUE}
set.seed(2)
chance <- CardDeck$new(chancedeck, verbose = FALSE)
community <- CardDeck$new(communitydeck, verbose = FALSE)
dice <- RandomDice$new()

player1 <- Player$new(verbose = FALSE)
player2 <- Player$new(verbose = FALSE)
monopoly <- SpaceTracker$new(verbose = FALSE)

for(g in 1:1000) {
  if(g %% 100 == 0) {
    cat("#### SIMULATING GAME", g, "##### \n")
  }
  for(i in 1:150){
    take_turn(player1, monopoly)
    take_turn(player2, monopoly)
  }
}

print(setNames(monopoly$counts, 1:40))
barplot(monopoly$counts)

results <- cbind(gameboard, tally = monopoly$counts)
results <- cbind(results, rel = monopoly$counts/sum(monopoly$counts))
print(results)

arrange(results, desc(tally))
```

